---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# LRTesteR Overview
<!-- badges: start -->
[![R build status](https://github.com/gmcmacran/LRTesteR/workflows/R-CMD-check/badge.svg)](https://github.com/gmcmacran/LRTesteR/actions)
[![CRAN status](https://www.r-pkg.org/badges/version/LRTesteR)](https://cran.r-project.org/package=LRTesteR)
<!-- badges: end -->

LRTesteR provides likelihood ratio tests and associated confidence intervals for many common distributions. All parametric tests and confidence intervals rely on the $\chi^2$ approximation even when exact sampling distributions are known. Parametric tests require a sample size of at least 50. Estimated asymptotic type I and type II error rates can be found [here](https://github.com/gmcmacran/TypeOneTypeTwoSim).

All functions match popular tests in R. If you are familiar with t.test and binom.test, you already know how to use these functions.

# Nonparametric Tests and Confidence Intervals

* Empirical Likelihood
  + mean
  
* Bootstrapping
  + any statistic
  
# Parametric Tests and Confidence Intervals
  
* Beta
  + shape 1
  + shape 2

* Binomial
  + p
  
* Exponential
  + rate
  
* Gamma
  + rate
  + scale
  + shape
  
* Gaussian
  + mu
  + variance
  
* Negative Binomial
  + p
  
* Poisson
  + lambda

* Cauchy
  + location
  + scale
  
* Inverse Gaussian
  + mean
  + shape
  + dispersion

# Example 1: Test lambda of a poisson distribution

To test lambda, simply call poisson_lambda_one_sample.

```{r example1Part1}
library(LRTesteR)

set.seed(1)
x <- rpois(n = 100, lambda = 1)
poisson_lambda_one_sample(x = x, lambda = 1, alternative = "two.sided")

```

# Example 2: Confidence Interval

To get a confidence interval, set the conf.level to the desired confidence. Below gets a two sided 90% confidence interval for scale from a Cauchy random variable.

```{r example2Part1}
set.seed(1)
x <- rcauchy(n = 100, location = 3, scale = 5)
cauchy_scale_one_sample(x = x, scale = 5, alternative = "two.sided", conf.level = .90)
```

Setting alternative to "less" gets a lower one sided interval. 
```{r example2Part2}
cauchy_scale_one_sample(x = x, scale = 5, alternative = "less", conf.level = .90)
```


Setting it to "greater" gets an upper one sided interval.
```{r example2Part3}
cauchy_scale_one_sample(x = x, scale = 5, alternative = "greater", conf.level = .90)
```

# Example 3: One-way Analysis
One-way ANOVA is generalized to all distributions. Here gamma random variables are created with different shapes. The one way test has a small p value and provides confidence intervals with 95% confidence for the whole set.

```{r example3Part1}
set.seed(1)
x <- c(rgamma(50, 1, 2), rgamma(50, 2, 2), rgamma(50, 3, 2))
fctr <- c(rep(1, 50), rep(2, 50), rep(3, 50))
fctr <- factor(fctr, levels = c("1", "2", "3"))
gamma_shape_one_way(x, fctr, .95)
```

# The $\chi^2$ approximation

As implemented, all functions depend on the $\chi^2$ approximation. To get a sense of performance, lets compare the likelihood method to the exact method. Here, X is normally distributed with mu equal to 3 and standard deviation equal to 2. The two intervals are similar.

```{r MDPart1}
set.seed(1)
x <- rnorm(n = 50, mean = 3, sd = 2)
exactTest <- t.test(x = x, mu = 2.5, alternative = "two.sided", conf.level = .95)
likelihoodTest <- gaussian_mu_one_sample(x = x, mu = 2.5, alternative = "two.sided", conf.level = .95)
as.numeric(exactTest$conf.int)
likelihoodTest$conf.int
```

Lets compare tests for variance. Again, confidence intervals are similar.
```{r MDPart3}
sigma2 <- 1.5^2 # Variance, not standard deviation.
exactTest <- EnvStats::varTest(x = x, sigma.squared = sigma2,  alternative = "two.sided", conf.level = .95)
likelihoodTest <- gaussian_variance_one_sample(x = x, sigma.squared = sigma2, alternative = "two.sided", conf.level = .95)
as.numeric(exactTest$conf.int)
likelihoodTest$conf.int
```

Changing to p for a binomial distribution, the confidence intervals are similar yet again. 
```{r MDPart4}
exactTest <- stats::binom.test(x = 10, n = 50,  p = .50, alternative = "two.sided", conf.level = .95)
likelihoodTest <- binomial_p_one_sample(x = 10, n = 50,  p = .50, alternative = "two.sided", conf.level = .95)
as.numeric(exactTest$conf.int)
likelihoodTest$conf.int
```

When exact methods are known, use them. The utility of the likelihood based approach is its generality. Many tests in this package don't have other well known options.
