---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# LRTesteR Overview
<!-- badges: start -->
[![R build status](https://github.com/gmcmacran/LRTesteR/workflows/R-CMD-check/badge.svg)](https://github.com/gmcmacran/LRTesteR/actions)
[![CRAN status](https://www.r-pkg.org/badges/version/LRTesteR)](https://cran.r-project.org/package=LRTesteR)
<!-- badges: end -->

LRTesteR implements the likelihood ratio test for many common distributions. All tests rely on the chi-square approximation even when exact sampling distributions are known. Tests require a sample size of at least 50. Estimated asymptotic type I and type II error rates can be found [here](https://github.com/gmcmacran/TypeOneTypeTwoSim).

# Implemented Tests

* beta
  + shape 1 test
  + shape 2 test

* binomial
  + p test
  
* exponential
  + rate test
  
* gamma
  + rate test
  + scale test
  + shape test
  
* gaussian
  + mu test
  + variance test
  
* negative binomial
  + p test
  
* poisson
  + lambda test

# Example 1: Test lambda of a poisson distribution

Lets create some data to work with.

```{r example1Part1}
library(LRTesteR)

set.seed(1)
x <- rpois(n = 100, lambda = 1)
```

To test lambda, simply call poisson_lambda_lr_test.
```{r example1Part2}
poisson_lambda_lr_test(x = x, lambda = 1, alternative = "two.sided")
```
Because we generated the data, we know the true value of lambda is one. The p value is above 05%.

# Example 2: Test rate of an exponential distribution

This time, lets run a test where the null is false. Below the true rate is 3 and the rejected null tests a rate of 1.

```{r example2Part1}
library(LRTesteR)

set.seed(1)
x <- rexp(n = 100, rate = 3)
exponentail_rate_lr_test(x = x, rate = 1, alternative = "two.sided")
```



