---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# LRTesteR

<!-- badges: start -->
[![R build status](https://github.com/gmcmacran/LRTesteR/workflows/R-CMD-check/badge.svg)](https://github.com/gmcmacran/LRTesteR/actions)
[![Codecov test coverage](https://codecov.io/gh/gmcmacran/LRTesteR/branch/main/graph/badge.svg)](https://app.codecov.io/gh/gmcmacran/LRTesteR?branch=main)
[![CRAN status](https://www.r-pkg.org/badges/version/LRTesteR)](https://cran.r-project.org/package=LRTesteR)
<!-- badges: end -->

LRTesteR implements the likelihood ratio test for many common distributions. All tests rely on chi-square approximation even when exact sampling distributions are known. Tests require a sample size of at least 50. Estimated asymptotic type I and type II error rates can be found [here](https://github.com/gmcmacran/TypeOneTypeTwoSim).

* beta
  + shape 1 test
  + shape 2 test

* binomial
  + p test
  
* exponential
  + rate test
  
* gamma
  + rate test
  + scale test
  + shape test
  
* gaussian
  + mu test
  + variance test
  
* geometric
  + p test
  
* negative binomial
  + p test
  
* poisson
  + lambda test

# Test mu of a gaussian distribution. 

```{r example}
library(LRTesteR)

# Null is true
set.seed(1)
x <- rnorm(100, 0, 1)
gaussian_mu_lr_test(x, 0, "two.sided")

# Null is false
set.seed(1)
x <- rnorm(100, 3, 1)
gaussian_mu_lr_test(x, 0, "greater")
```



