---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# LRTesteR Overview
<!-- badges: start -->
[![R-CMD-check](https://github.com/gmcmacran/LRTesteR/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/gmcmacran/LRTesteR/actions/workflows/R-CMD-check.yaml)
[![CRAN status](https://www.r-pkg.org/badges/version/LRTesteR)](https://cran.r-project.org/package=LRTesteR)
<!-- badges: end -->

LRTesteR provides 42 hypothesis tests and confidence intervals based on the likelihood ratio statistic. Of these, 36 are parametric tests about a specific distribution and 6 are nonparametric tests based on the empirical likelihood. Every test comes in two forms: a one sample version and an ANOVA style version.


# Example 1: Test lambda of a poisson distribution

To test lambda, simply call poisson_lambda_test.

```{r example1Part1}
library(LRTesteR)

set.seed(1)
x <- rpois(n = 100, lambda = 1)
poisson_lambda_test(x = x, lambda = 1, alternative = "two.sided")
```

# Example 2: Confidence Interval

To get a confidence interval, set conf.level to the desired confidence. The example below gets a two sided 90% confidence interval for the scale of a Cauchy random variable.

```{r example2Part1}
set.seed(1)
x <- rcauchy(n = 100, location = 3, scale = 5)
cauchy_scale_test(x = x, scale = 5, alternative = "two.sided", conf.level = .90)
```

Setting alternative to "less" gets a one sided interval bounded above.
```{r example2Part2}
cauchy_scale_test(x = x, scale = 5, alternative = "less", conf.level = .90)
```


Setting it to "greater" gets a one sided interval bounded below.
```{r example2Part3}
cauchy_scale_test(x = x, scale = 5, alternative = "greater", conf.level = .90)
```

# Example 3: One-way Analysis
One-way ANOVA is generalized to all distributions. Here gamma random variables are created with different shapes. The test returns a small p value and a set of confidence intervals that have 95% confidence as a group.

```{r example3Part1}
set.seed(1)
x <- c(rgamma(n = 50, shape = 1, rate = 2), rgamma(n = 50, shape = 2, rate = 2), rgamma(n = 50, shape = 3, rate = 2))
fctr <- c(rep(1, 50), rep(2, 50), rep(3, 50))
fctr <- factor(fctr, levels = c("1", "2", "3"))
gamma_shape_one_way_test(x = x, fctr = fctr, conf.level = .95)
```

# Example 4: Empirical Likelihood
The empirical likelihood tests do not require any distributional assumptions and work with less data.

```{r example4Part1}
set.seed(1)
x <- rnorm(n = 25, mean = 1, sd = 1)
empirical_mu_test(x = x, mu = 1, alternative = "two.sided")
```

See the Nonparametric Tests vignette for more detail on the empirical likelihood.

# The &chi;<sup>2</sup> approximation

As implemented, all functions depend on the asymptotic &chi;<sup>2</sup> approximation. To get a sense of how accurate this approximation is for large samples, the likelihood tests are compared to the exact tests below.

First, X is normally distributed with &mu; equal to 3 and standard deviation equal to 2. The two intervals for &mu; are similar.

```{r MDPart1}
set.seed(1)
x <- rnorm(n = 500, mean = 3, sd = 2)
exactTest <- t.test(x = x, mu = 2.5, alternative = "two.sided", conf.level = .95)
likelihoodTest <- gaussian_mu_test(x = x, mu = 2.5, alternative = "two.sided", conf.level = .95)
as.numeric(exactTest$conf.int)
likelihoodTest$conf.int
```

The confidence intervals for variance are similar as well.
```{r MDPart2}
set.seed(1)
x <- rnorm(n = 500, mean = 3, sd = 2)
sigma2 <- 1.5^2 # Variance, not standard deviation.
exactTest <- EnvStats::varTest(x = x, sigma.squared = sigma2, alternative = "two.sided", conf.level = .95)
likelihoodTest <- gaussian_variance_test(x = x, sigma.squared = sigma2, alternative = "two.sided", conf.level = .95)
as.numeric(exactTest$conf.int)
likelihoodTest$conf.int
```

Switching to p of a binomial random variable, the confidence intervals are similar yet again.
```{r MDPart3}
exactTest <- stats::binom.test(x = 250, n = 500, p = .50, alternative = "two.sided", conf.level = .95)
likelihoodTest <- binomial_p_test(x = 250, n = 500, p = .50, alternative = "two.sided", conf.level = .95)
as.numeric(exactTest$conf.int)
likelihoodTest$conf.int
```

When sample size is small, the similarity decreases. When exact methods are available, they are the better option. The utility of the likelihood based approach is its generality: this package provides hypothesis tests and confidence intervals for distributions that otherwise would not have them.

Estimated asymptotic type I and type II error rates can be found [here](https://github.com/gmcmacran/TypeOneTypeTwoSim). 
