---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# MLTesteR

<!-- badges: start -->
[![R build status](https://github.com/gmcmacran/MLTesteR/workflows/R-CMD-check/badge.svg)](https://github.com/gmcmacran/MLTesteR/actions)
[![Codecov test coverage](https://codecov.io/gh/gmcmacran/MLTesteR/branch/master/graph/badge.svg)](https://codecov.io/gh/gmcmacran/MLTesteR?branch=master)
[![CRAN status](https://www.r-pkg.org/badges/version/MLTesteR)](https://cran.r-project.org/package=MLTesteR)
<!-- badges: end -->

MLTester is a work in progress. The goal is to implement the likelihood ratio, wald and score tests for all common families. Currently a few likelihood ratio tests are implemented. A thorough evaluation (type 1 error rate, publication comparisons, thorough test creation, etc) have not been done yet.

Package is mostly of theoretical interest. Exact tests are implemented in other R packages. These tests are all approximations.

## Installation

And the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("gmcmacran/MLTesteR")
```
## Example

This is a basic example which shows how to test the mean of a Gaussian distribution with the likelihood ratio test:

```{r example}
library(MLTesteR)

# Null is true
set.seed(1)
x <- rnorm(100, 0, 1)
gaussian_mu_lr_test(x, 0, "two.sided")

# Null is false
set.seed(1)
x <- rnorm(100, 3, 1)
gaussian_mu_lr_test(x, 0, "greater")
```



