---
title: "Analysis of Deviance"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Analysis of Deviance}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Classic One-Way ANOVA

As an introduction, lets start with one way ANOVA. Here three random variables following a normal distribution with a common standard deviation are created. For this test, the null hypothesis is

$$ H_{0}: \mu_0 = \mu_1 = \mu_2  $$

```{r part1}
library(LRTesteR)

set.seed(123)
x <- c(
  rnorm(n = 50, mean = 1, sd = 1),
  rnorm(n = 50, mean = 3, sd = 1),
  rnorm(n = 50, mean = 5, sd = 1)
)
fctr <- c(rep(1, 50), rep(2, 50), rep(3, 50))
fctr <- factor(fctr, levels = c("1", "2", "3"))
gaussian_mu_one_way(x = x, fctr = fctr, conf.level = 0.95)
```
Statistically, there are number of things important to note:

* The statistic is the log likelihood statistic, not the F statistic.
* The p value is based on the chi-squared distribution, not the F distribution.
* The individual confidence intervals use the Bonferroni correction.

For a normal distribution, the above is not optimal. The F statistic is exact. The log likelihood statistic is an asymptotic approximation. The Bonferroni correction will lead to the widest confidence intervals compared to any other correction technique. 

The principal advantage of the likelihood approach is it can be applied nonparametrically and to other distributions.

# Nonparametric One-Way ANOVA

One-way analysis without assuming the data are normally distributed.

```{r part2}
empirical_mu_one_way(x = x, fctr = fctr, conf.level = 0.95)
```

# Nonparametric Equality Of Variance
$$ H_{0}: \sigma^2_0 = \sigma^2_1 = \sigma^2_2  $$
```{r part4}
empirical_variance_one_way(x = x, fctr = fctr, conf.level = 0.95)
```

# Cauchy Random Variables
Here two random variables following a Cauchy distribution with a common location and different scales are created. For this test, the null hypothesis is 

$$ H_{0}: \gamma_0 = \gamma_1  $$

```{r part5}
set.seed(1)
x <- c(rcauchy(n = 50, location = 2, scale = 1), rcauchy(n = 50, location = 2, scale = 3))
fctr <- c(rep(1, 50), rep(2, 50))
fctr <- factor(fctr, levels = c("1", "2"))
cauchy_scale_one_way(x = x, fctr = fctr, conf.level = 0.95)
```



# Poisson Random Variables
Here three poisson random variables with different lambdas are created. The null hypothesis is 

$$ H_{0}: \lambda_0 = \lambda_1 = \lambda_2  $$

```{r part6}
set.seed(1)
x <- c(rpois(n = 50, lambda = 1), rpois(n = 50, lambda = 2), rpois(n = 50, lambda = 3))
fctr <- c(rep(1, 50), rep(2, 50), rep(3, 50))
fctr <- factor(fctr, levels = c("1", "2", "3"))
poisson_lambda_one_way(x = x, fctr = fctr, conf.level = 0.95)
```
