---
title: "Nonparametric Tests"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Nonparametric Tests}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Empirical Likelihood Tests

Empirical likelihood is a nonparametric method of inference based on a data driven likelihood ratio function. Like the bootstrap and jackknife, empirical likelihood inference does not require us to specify a family of distributions for the data. Like parametric likelihood methods, empirical likelihood makes an automatic determination of the shape of confidence regions [...] and has very favorable asymptotic power properties. It can be thought of as a bootstrap that does not resample and as a likelihood without parametric assumptions. (Owen 2001)

The following demonstrates the empirical likelihood approach produces similar p values and similar confidence internals in optimal conditions for the parametric method. Even when the parametric method's assumptions are true, the assumption offers little advantage. 

## Tests For Mu

### Gaussian Distribution

```{r part1}
library(LRTesteR)
library(statmod)

set.seed(1)
x <- rnorm(50)

gaussian_mu_test(x = x, mu = 0, alternative = "two.sided")
```

```{r part2}
empirical_mu_test(x = x, mu = 0, alternative = "two.sided")
```

### Inverse Gaussian Distribution

```{r part3}
set.seed(1)
x <- rinvgauss(50)

inverse_gaussian_mu_test(x = x, mu = 1, alternative = "two.sided")
```
The one way test does not require nuisance parameters or sample size to be equal.

```{r part4}
empirical_mu_test(x = x, mu = 1, alternative = "two.sided")
```


## Tests For Variance

### Gaussian Distribution
```{r part6}
set.seed(1)
x <- rnorm(50)

gaussian_variance_test(x = x, sigma.squared = 1, alternative = "two.sided")
```
```{r part7}
empirical_variance_test(x = x, sigma.squared = 1, alternative = "two.sided")
```

## One Way Tests For Mu

### Gaussian Distribution
```{r part8}
set.seed(1)
x <- rnorm(100)
fctr <- c(rep(1, 50), rep(2, 50))
fctr <- factor(fctr, levels = c("1", "2"))

gaussian_mu_one_way_test(x = x, fctr = fctr, conf.level = .95)
```

```{r part9}
empirical_mu_one_way_test(x = x, fctr = fctr, conf.level = .95)
```

### Inverse Gaussian Distribution
```{r part10}
set.seed(1)
x <- rinvgauss(100)
fctr <- c(rep(1, 50), rep(2, 50))
fctr <- factor(fctr, levels = c("1", "2"))

inverse_gaussian_mu_one_way_test(x = x, fctr = fctr, conf.level = .95)
```

```{r part11}
empirical_mu_one_way_test(x = x, fctr = fctr, conf.level = .95)
```

## One Way Tests For Variance

### Gaussian Distribution
```{r part12}
set.seed(1)
x <- rnorm(100)
fctr <- c(rep(1, 50), rep(2, 50))
fctr <- factor(fctr, levels = c("1", "2"))

gaussian_variance_one_way_test(x = x, fctr = fctr, conf.level = .95)
```

```{r part13}
empirical_variance_one_way_test(x = x, fctr = fctr, conf.level = .95)
```
